---
title: "Lab 4"
author: Camille Heathcock
format: 
  html:
    self-contained: true
editor: visual
---

```{r}
#| echo: False 
#| output: False
library(tidyverse)
avocado_data <- read_csv(here::here("datasets", "avocado.csv"))
```

## Set-up

#### 1) Dataset Description

The data set contains information about avocados prices. The idea behind the data collection was to find a region in which avocado prices are low enough to allow millennials to buy their fix of avocado toast and also be able to buy a home.

#### 2) Dataset Cleaning

```{r}
avocado_data <- avocado_data |> 
  rename("xl_avocados" = `4770`, 
         "large_avocados" = `4225`, 
         "small_avocados" = `4046`)
```

```{r}
avocado_totalregions <- avocado_data |> 
  distinct(region)

states_data <- data.frame("state" = c("California", "SouthCarolina"))

regions_data <- data.frame("regions" = c("GreatLakes", "Midsouth", "Northeast", "NorthernNewEngland", "Plains", "SouthCentral", "Southeast", "West", "WestTexNewMexico"))

totalus_data <- data.frame("totalus" = c("TotalUS"))

avocado_states <- avocado_data |> 
  semi_join(states_data, by = c(region = "state"))

avocado_regions <- avocado_data |> 
  semi_join(regions_data, by = c(region = "regions"))

avocado_metro_regions <- avocado_data |> 
  anti_join(states_data, by = c("region" = "state")) |> 
  anti_join(regions_data, by = c("region" = "regions")) |> 
  anti_join(totalus_data, by = c("region" = "totalus"))
```

## Exercises

#### 3) Which major region sold the most organic, small Hass avocados in 2017?

```{r}
avocado_regions |> 
  filter(year == 2017, 
         type == "organic") |>
  group_by(region) |> 
  summarize(mean_sum = mean(small_avocados)) |> 
  slice_max(mean_sum)
```

The region that sold the most organic, small Hass avocados in 2017 was the West with 35,287 sales.

#### 4) Use the separate() function to split the Date variable into year, month, and day. In which month is the highest volume of avocado sales?

```{r}
avocado_data |> 
  separate(col = Date, sep = "-", into = c("Year", "Month", "Day")) |> 
  group_by(Month) |> 
  summarize(mean_vol = mean(`Total Volume`)) |> 
  slice_max(mean_vol)
```

The month with the highest volume of avocado sales is February with 1,018,825 avocado sales.

#### 5) Which metro area regions sold the most total avocados? Plot side-by-side boxplots of total volume for only the five regions with the highest averages for the Total Volume variable.

**Regions with most avocado sales:**

```{r}
topfive <- avocado_metro_regions |>  
  group_by(region) |> 
  summarize(total_volume = mean(`Total Volume`)) |> 
  arrange(desc(total_volume)) |> 
  slice_head(n = 5)
```

The metro region that sold the most avocados is Los Angeles, with 1,502,653 total avocado sales. This is followed by New York, Dallas/ Ft. Worth, Houston, and Phoenix / Tucson.

**Boxplot:**

```{r}
top_five <- avocado_metro_regions |> 
  semi_join(topfive, by = c(region = "region"))

top_five |> 
  ggplot(mapping = aes(x = `Total Volume`,
                     y = region)) + 
  geom_boxplot()
```

## Reshaping

#### 6) In which regions is the price of organic versus conventional avocados most different? Support your answer with a few summary statistics and a visualization.

```{r}
california_baby <- data.frame(region = c("LosAngeles", "SanDiego", "Sacramento", "SanFrancisco"))

california_regions <- avocado_metro_regions |> 
  semi_join(california_baby, by = c(region = "region"))

california_regions |> 
  group_by(region, type) |> 
  summarize(mean_price = mean(AveragePrice), .groups = "keep") |> 
  pivot_wider(names_from = type, values_from = mean_price) |> 
  mutate(diff = conventional - organic)

#visualization 
california_regions |>
  ggplot(mapping = aes(y = `Total Volume`, 
                     x = region, 
                     color = type)
       ) + 
  geom_jitter()

```

The region where the price of organic and conventional avocados is the most different is San Francisco, with organic avocados being \$0.81 more than conventional, on average.

#### 7) Recreate the plot

```{r}
#| output: False 
ca_regions_long <- california_regions |>
  pivot_longer(cols = small_avocados:xl_avocados, 
             names_to = "Size", 
             values_to = "Sales_Volume") |> 
  group_by(region, Size, type) |> 
  summarize(mean_sales = mean(Sales_Volume))
```

```{r}
ca_regions_long |>
  ggplot(mapping = aes(y = mean_sales, 
                     x = region, 
                     fill = Size)) + 
  geom_col(position = "fill") + 
  facet_wrap(~ type) + 
  xlab("Region of CA") +
  ylab("Proportion of Avocados Sold") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))
```
