---
title: "Lab 5"
author: Camille Heathcock
format: 
  html:
    self-contained: true
editor: visual
---

```{r}
#| echo: False 
#| output: False
library(tidyverse)
library(here)
surveys <- read_csv(here::here("datasets", "surveys.csv"))
```

**1. Reorder the box plots so the weights go in descending order**

```{r}
#lab 2 plot 
surveys |> 
  ggplot(mapping = aes(y = fct_reorder(species, 
                                       weight, 
                                       .desc = TRUE), 
                       x = weight)) + 
  geom_boxplot() + 
  labs(x = "Weight (g)",
       y = "Species of Animal")  
```

**2. Now that you've reordered, look at your y-axis label. That's long! Let's take Will Chase's advice and move the y-axis label to the top of the plot.**

```{r}
surveys |> 
  ggplot(mapping = aes(y = fct_reorder(species, weight, .desc = TRUE), 
                       x = weight)) + 
  geom_boxplot() + 
  labs(x = "Weight (g)", 
       y = " ") +
  ggtitle("Species of Rodent by Weight")
```

## Time Series Plot

**1. Create a visualization how weights vary for each genus over the duration of the study.**

```{r}
surveys |> 
  ggplot(aes(x = year, 
             y = weight)) + 
  geom_line() + 
  facet_wrap( ~ genus) +
  labs(title = "Genus Weight Variation over Duration of Study", 
       x = "Year", 
       y = "Weight (g)")
```

**2. Calculate and plot the mean weight for each year (for each genus).**

```{r}
#| message: False
mean_yearly_weight <- surveys |> 
  group_by(year, genus) |> 
  summarize(mean_weight = mean(weight))

surveys_exp <- full_join(surveys, mean_yearly_weight)

surveys_exp |> 
  ggplot(aes(x = year, 
             y = mean_weight, 
             color = genus)) +
  geom_line() + 
  labs(title = "Genus Mean Weight Variation over Duration of Study", 
       x = "Year", y = "Mean Weight (g)") 
```

**3. Reorder the lines so the weights go in descending order.**

```{r}
surveys_exp |> 
  ggplot(aes(x = year, 
             y = mean_weight, 
             color = fct_reorder(genus, mean_weight, .desc = TRUE))) +
  geom_line() + 
  labs(title = "Genus Mean Weight Variation over Duration of Study", 
       x = "Year", 
       y = "Mean Weight (g)")
```

**4. Now that you've reordered the lines, look at your legend's label. That's terrible! Let's give the legend a new title *and* move the information from the y-axis to the plot title.**

```{r}
surveys_exp |> 
  ggplot(aes(x = year, 
             y = mean_weight, 
             color = fct_reorder(genus, mean_weight, .desc = TRUE))) +
  geom_line() + 
  labs(title = "Genus Mean Weight (g) Variation over Duration of Study", 
       x = "", 
       y = "", 
       color = "Genus")
```

## **Captures over the Week**

**1. Create a visualization of the number of rodents captured each day of the week.**

```{r}
surveys_daily <- surveys |> 
  group_by(day_of_week) |> 
  drop_na(day_of_week) |> 
  count(day_of_week)
  
surveys |> 
  drop_na(day_of_week) |>
  ggplot(mapping = aes(x = fct_relevel(day_of_week))) + 
  geom_bar(fill = "purple") + 
  labs(x = "Day of the Week", 
       y = "", 
       title = "Number of Rodents Captured Each Day of the Week")

```

**2. Change the order of the day of the week to go Monday through Sunday.**

```{r}
surveys |> 
  drop_na(day_of_week) |>
  ggplot(mapping = aes(x = fct_relevel(day_of_week, 
                                     c("Mon", 
                                       "Tue", 
                                       "Wed", 
                                       "Thu", 
                                       "Fri", 
                                       "Sat", 
                                       "Sun")))) + 
  geom_bar(fill = "pink") + 
  labs(x = "Day of the Week", 
       y = "", 
       title = "Number of Rodents Captured Each Day of the Week")
```

**3. Collapse Monday through Friday into a `"Weekday"` level, and collapse Saturday and Sunday into a `"Weekend"` level. Plot the number of rodents captured between the two groups.**

```{r}
surveys <- surveys |> 
  drop_na(day_of_week) |>
  mutate(time_of_week = fct_collapse(day_of_week,
               Weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"), 
               Weekend = c("Sat", "Sun")))

ggplot(data = surveys, mapping = aes(x = time_of_week)) + 
  geom_bar(fill = "orange") +
  labs(x = "Time of Week", 
       y = "", 
       title = "Number of Rodents Captured During Weekdays vs Weekends")
```
